---
title: "PRBpower - Compute The Power of a prospective-retrospetive Biomarker Study with a Time To Event Outcome"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PRBpower}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PRBpower)
library(survival)
library(dplyr)
library(ggplot2)
```

```{r}
# user supplied input ========================================================
ppos    <- 0.20        # proportion of subjects expected M+ in trial population 
HR      <- c(1.0, 0.5) # posited BM haz ratios M+ : M- for std and experimental treatment groups. 
alpha   <- 0.05        # Type I error allocated to the biomarker study.
nsim    <- 1000        # number of simulated trials to generate. 0 = no simulations. 
# set.seed(20231003)


data(parentTrial)      # pull in the data from the parent clinical trial
# perform data transformations to have characteristics required for PRBpower
# Specifically the parent data frame should have one records for each subject 
# and at least the following 3 variables:
#   nrx (treatment code; 0=standard treatment, 1=experimental(targeted) treatment).
#   survtime (time at risk of the study event).
#   survstat (indicator for whether subjected evented: 0=censored, 1=evented)
dfp <- parentTrial %>% select(nrx, survtime, survstat) %>%
        arrange(., nrx,survtime)
dfp$nrx      <- ifelse(dfp$nrx==1, 0, 1)          # recode df$rx (1,2) -> (0,1)
dfp$survstat <- ifelse((dfp$survstat == 1), 1, 0) # recode censor indicator. 1=event; 0=censor
dfp$rx       <- factor(dfp$nrx, levels=c(0,1), labels=c("std", "Exp"),ordered=TRUE) # std is ref grp.

utils::head(dfp)

# paramsCheck(ppos=ppos, HR=HR, nsim=nsim, alpha=alpha, dfp=dfp)
  
# calculate Kaplan-Meier estimates
kmdef <- Surv(dfp$survtime, dfp$survstat == 1)
kmfit <- survfit(kmdef ~ dfp$rx ) 
  
dfm <- dissect_km (ppos, HR, kmfit) # separate Trtmt Group KM into M- & M+
  
dfm <- ppos_calc (dfm, ppos) # compute probability of being M+ for each event/censor time 
  
# join parent trial data, df, to the estimated probabilities each subject is M+ 
dfx <- join_parent (dfp, dfm)

utils::head(dfx)
  
  # summarize number of observed events by treatment group in the parent trial. Also,  
  # summarize expected number of events by treatment and Marker status under alt hypothesis. 
  eventsTable.obs <- eventsTable_obs(dfx)
  eventsTable.obs
  eventsTable.exp <- eventsTable_exp (dfx) 
  eventsTable.exp
  
  power.PG <- power_PG(EE=eventsTable.exp, HR=HR, alpha=alpha)
  power.PG
  
  # if simulations are requested then do them 
  if (nsim > 0) {
    df.sim <- sim_gen (dfx, nsim)
    power.sim <- sum(df.sim[ , "p.nrx.BM"] <= alpha)/nsim
    print(power.sim)
    RH.tab <- calc_RH(df.sim)
    RH.tab
  } else {
    print(" === No simulations were requested")
  }

  plot_PRB (dfx=dfx, plotType=3, ppos=ppos, HR=HR )
  
